---
title: "Lecture 2"
author: "Chiong"
date: "8/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) 
library(stargazer)
library(RPostgres)
library(logKDE)
```

### Probability integral transformation

Draw $n=1000$ random uniform numbers: $(u_{1},\dots,u_{n})$ from $U[0,1]$. Plot a histpgram of these random uniform numbers.


```{r}
u <- runif(1000,min=0, max=1)
hist(u)
```

### Inverse CDF

Our target cdf is: $F_{X}(x) = 1- \exp(-x)$, which is the exponential distribution with $\lambda = 1$. The general cdf is $F_{X}(x) = 1- \exp(-\lambda x)$, with the pdf $f_{X}(x) = \lambda \exp(-\lambda x)$.

The inverse of this cdf is: $\log(\frac{1}{1-u_{i}})$

```{r}
x <- log(1/(1-u))
hist(x)
hist(x,prob=TRUE)
lines(density(x))
```

### Comparison with the R's exponential distribution

```{r}
r <- rexp(1000, rate = 1)
hist(r)
hist(r, prob=TRUE)
lines(density(r)) 
lines(density(x),col="blue", lwd=2,lty="dotted")
```

### Comparison with the theoretical pdf


```{r}
z <- seq(from = 0, to = 5, by = 0.05)
plot(z,exp(-z))
lines(density(x),col="blue", lwd=2,lty="dotted")
```

### Log-transformation

```{r}
hist(log(x), prob=TRUE)
lines(density(log(x)))
```
The pdf of $Y= log(X)$ is?

$F_{Y}(y) = P(Y<=y) = P(log(X)<=y) = P(X<=e^{y}) = 1-e^{-e^{y}}$.

Therefore the pdf of $Y$ is $f_{Y}(y) = e^{y}e^{-e^{y}}$

```{r}
z <- seq(from = -5, to = 2, by = 0.05)
plot(z,exp(z)*exp(-exp(z)))
lines(density(log(x)))
```

