---
title: "PS3"
author: "Chiong"
date: "12/9/2019"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr) 
library(tidyverse)
```

#Solutions to Problem Set 3

##Question 4

```{r}
a0 <- 2 #null hypothesis
n <- 100 #sample size
s <- 2000 #number of simulations

#generate data
x <- matrix(rbeta(n*s, a0,1), nrow = n, ncol=s) 
#sample product for each column
m <- apply(x,2,prod) 

#Likelihood ratio test statistics across simulations
lr <- ((-a0*log(m)/n)^n)*m^(a0 + n/log(m))

#the exact critical value for a size-0.05 likelihood ratio test is:
c <- quantile(lr, c(0.05))

#compare to the asymptotic critical value
quantile(lr, c(0.05)) - exp(-qchisq(0.95,df=1)/2)
```

## For n large, verify that the Likelihood Ratio Test statistic converges in distribution to the Chi-Squared distribution.

```{r}
hist(-2*log(lr),breaks = 20, probability = TRUE) 

#comparing histogram with the pdf of chi-squared distribution
z <- seq(min(-2*log(lr)),max(-2*log(lr)),0.01)
lines(z,dchisq(z, 1),col="blue", lwd=4,lty="dotted")
```

## Using simulation to derive the power function for the Likelihood Ratio test

```{r}
a0<- 2
#support of the power function
support <- seq(from=0.1, to=3, length.out = 40) 

#stores the values of the power function
B <- vector('numeric')

for (a in support) {
  
  #generate data
  x <- matrix(rbeta(n*s, a,1), nrow = n, ncol=s) 
  #sample product for each column
  m <- apply(x,2,prod) 

  #Likelihood ratio test statistics across simulations
  lr <- ((-a0*log(m)/n)^n)*m^(a0 + n/log(m))

  #sum(lr<0.1465)/s is the fraction of simulations in which the null is rejected 
  B <- rbind(B,sum(lr<0.1465)/s)
  
}

plot(support, B, ylim = c(0,1),type = "o",col ="blue",ylab="Power Function",xlab = "alpha",lwd=1.5,panel.first=grid(),xaxp  = c(0, 3, 15),yaxp  = c(0, 1, 20))
```

## Using simulation to derive the power function for the Wald test

```{r}
a0 <- 2

#support of the power function
support <- seq(from=0.1, to=3, length.out = 40) 

#stores the values of the power function
B2 <- vector('numeric')

for (a in support) {
  
  #generate data
  x <- matrix(rbeta(n*s, a,1), nrow = n, ncol=s) 
  #sample mean for each column
  m <- apply(x,2,mean) 

  #Wald test statistics across simulations
  w <- sqrt(n)*(m/(1-m) - a0)/sqrt((a0*(1+a0)^2)/(2+a0))

  #sum(abs(w) > 1.96)/s) is the fraction of simulations in which the null is rejected 
  B2 <- rbind(B2,sum(abs(w) > 1.96)/s)
  
}

plot(support, B, ylim = c(0,1),type = "o",col ="blue",ylab="Power Function",xlab = "alpha",lwd=1,panel.first=grid(),xaxp  = c(0, 3, 15),yaxp  = c(0, 1, 20))

lines(support, B2, ylim = c(0,1),type = "o",col ="red",ylab="Power Function",xlab = "alpha",lwd=1,panel.first=grid(),xaxp  = c(0, 3, 15),yaxp  = c(0, 1, 20))
```
## Now compare the power function using different values of n

```{r}
#support of the power function
support <- seq(from=0.1, to=3, length.out = 40) 

#stores the values of the power function
B1 <- vector('numeric')
B2 <- vector('numeric')
 
s <- 2000
n1 <- 20
n2 <- 100

for (a in support) {
  
  #generate data
  x1 <- matrix(rbeta(n1*s, a,1), nrow = n1, ncol=s) 
  x2 <- matrix(rbeta(n2*s, a,1), nrow = n2, ncol=s) 
  #sample mean for each column
  m1 <- apply(x1,2,mean) 
  m2 <- apply(x2,2,mean) 

  #Wald test statistics across simulations
  w1 <- sqrt(n1)*(m1/(1-m1) - a0)/sqrt((a0*(1+a0)^2)/(2+a0))
  w2 <- sqrt(n2)*(m2/(1-m2) - a0)/sqrt((a0*(1+a0)^2)/(2+a0))

  #sum(abs(w1) > 1.96)/s) is the fraction of simulations in which the null is rejected 
  B1 <- rbind(B1,sum(abs(w1) > 1.96)/s)
  B2 <- rbind(B2,sum(abs(w2) > 1.96)/s)
  
}

plot(support, B2, ylim = c(0,1),type = "o",col ="blue",ylab="Power Function",xlab = "alpha",lwd=1,panel.first=grid(),xaxp  = c(0, 3, 15),yaxp  = c(0, 1, 20))
lines(support, B1, ylim = c(0,1),type = "o",col ="red",ylab="Power Function",xlab = "alpha",lwd=1,panel.first=grid(),xaxp  = c(0, 3, 15),yaxp  = c(0, 1, 20))

```

# Question 5. Uniform distribution
## Using simulation to derive the power function for the Likelihood Ratio test

```{r}
#support of the power function
support <- seq(from=0, to=4, length.out = 40) 

#stores the values of the power function
B <- vector('numeric')
n <- 50
s <- 1000
theta0 <- 2

for (u in support) {
  
  #generate data
  x <- matrix(runif(n*s, 0,u), nrow = n, ncol=s) 
  #sample max for each column
  m <- apply(x,2,max) 

  B <- rbind(B,sum(m>theta0)/s)
  
}

plot(support, B, ylim = c(0,1),type = "o",col ="blue",ylab="Power Function",xlab = "upper bound",lwd=1.5,panel.first=grid(),xaxp  = c(0, 4, 10),yaxp  = c(0, 1, 20))
```

## Using simulation to derive the power function for the Wald test

```{r}
#support of the power function
support <- seq(from=0, to=4, length.out = 40) 

#stores the values of the power function
B2 <- vector('numeric')
n <- 50
s <- 1000
theta0 <- 2

for (u in support) {
  
  #generate data
  x <- matrix(runif(n*s, 0,u), nrow = n, ncol=s) 
  #sample max for each column
  m <- apply(x,2,mean) 

  w <- sqrt(n)*(2*m - theta0)/sqrt((theta0^2)/3)
  B2 <- rbind(B2,sum(w>1.64)/s)
  
}

plot(support, B, ylim = c(0,1),type = "o",col ="blue",ylab="Power Function",xlab = "upper bound",lwd=1.5,panel.first=grid(),xaxp  = c(0, 4, 10),yaxp  = c(0, 1, 20))
lines(support, B2, ylim = c(0,1),type = "o",col ="red",ylab="Power Function",xlab = "upper bound",lwd=1.5,panel.first=grid(),xaxp  = c(0, 4, 10),yaxp  = c(0, 1, 20))
```

