---
title: "Bootstrapping"
author: "Chiong"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr) 
library(tidyverse)
```

## R Markdown

Setup. Generate data

```{r}
n <- 500 #sample size
ld <- 0.5 #rate parameter
x <- rexp(n,ld) #this is our sample
```

##  Bootstrapping

```{r}
B <- 1000 #number of bootstrapped samples
Xb <- matrix(0, nrow = n, ncol = 0) #collects all bootstrapped samples
for (b in seq(1,B)) {
  Xb <- cbind(Xb,sample(x, n, replace = TRUE))
}
```

## Compute statistic of interest

```{r}
#define a function that computes the maximum likelihood estimator of variance 
svar <- function(X) {
  return(mean(X^2) - mean(X)^2)
}
t_boostrapped <- apply(Xb,2,svar)
hist(t_boostrapped,breaks = 15)
```

### Because we know the DGP, we can use monte carlo simulation to obtain the actual sampling distribution 

```{r}
s <- 1000  #number of simulations
X <- matrix(rexp(n*s,ld),nrow = n,ncol = s)
t_actual <- apply(X,2,svar)
hist(t_actual,breaks = 15)
```

### Compare the sampling distribution obtained using bootstrapping versus the actual sampling variance

```{r}
plot(density(t_boostrapped),col="red",main = "")
lines(density(t_actual),lty=2,lwd = 2,col="blue")
legend("topright", legend=c("Bootstrapped sampling distribution","Actual sampling distribution"),col=c("red", "blue"), lty=2, cex=0.8)
```


### Compare the sampling variance obtained using bootstrapping versus the actual sampling variance

```{r}
print(var(t_boostrapped))

print(var(t_actual))
```

